---
title: "Forward selection on conditional analysis results"
author: "Tayden Li"
output:
  html_document:
    keep_md: yes
---

Forward selectios allows us to arrive at a list of relatively independent significant cell types after conditional analyses. Forward selection criteria can be found on the [FUMA website](https://fuma.ctglab.nl/tutorial#celltype).

### Load libraries
```{r, message=F}
library(dplyr)
library(readr)
library(tidyr)
filename_cond <- "SCZ/Siletti_l2_joint_spe_sig-only_SCZ_2022.gsa.out"
filename_marg <- "SCZ/Siletti_l2_conti-spe_SCZ_2022.gsa.out"
```

### Read marginal and conditional results
```{r, message=F}
df_cond <- read_table(filename_cond, skip=4) %>%  # skip headers (3-4 lines)
  subset(select=-c(TYPE, NGENES, BETA, BETA_STD, SE)) %>% 
  rename(P_cond = P)
df_marg <- read_table(filename_marg, skip=4) %>%  # skip headers (3-4 lines)
  subset(select=-c(TYPE, NGENES, BETA, BETA_STD, SE)) %>% 
  rename(P_marg = P) %>% 
  arrange(P_marg)
n_sig <- ceiling(sqrt(nrow(df_cond)))
n_clusters <- nrow(df_marg)
```

### Combine the marginal and conditional results 
```{r}
df_comb <- left_join(df_cond, df_marg, by = "VARIABLE")
df_comb <- df_comb %>% 
  mutate(VarCode = rep(c("a", "b"),times=nrow(df_comb)/2)) %>%
  pivot_wider(names_from = VarCode, values_from=c(VARIABLE, P_cond, P_marg)) %>%
  mutate(PS_a=log10(P_cond_a)/log10(P_marg_a), PS_b=log10(P_cond_b)/log10(P_marg_b))
# assert order
stopifnot(df_comb$P_marg_a <= df_comb$P_marg_b)
```

### Reverse the order of forward selection if the criterion below is reached
Note that the reversed cell types are discarded here because they will not end up in the list of independent cell types according to our forward selection scheme.
```{r}
select_list <- df_marg$VARIABLE[1:n_sig]
reverse_list <- df_comb$VARIABLE_a[df_comb$PS_a < 0.2 & df_comb$PS_b >= 0.2]
select_list <- select_list[select_list %in% reverse_list == F][-1]
```

### Conduct forward selection
```{r, message=F}
# The independent list starts with the most marginally significant cell type
indep_list <- c(toString(df_marg[1,1])) 
for (cur_var in select_list) {
  for (indep_var in indep_list) {
    # drop if the condition indicates so with any cell types in indep_list
    df_cur = df_comb[df_comb$VARIABLE_a==indep_var & df_comb$VARIABLE_b==cur_var,]
    if ((df_cur$PS_a >= 0.8 & df_cur$PS_b >= 0.8) | 
        (df_cur$PS_a >= 0.5 & df_cur$PS_b >= 0.5 & df_cur$P_cond_b < 0.05)) {
      if (tail(indep_list,1) == indep_var) indep_list <- c(indep_list, cur_var)
    } else break
  }
}
cat(indep_list)
```
